---
title: "alcohol-related-admissions"
author: "D Blana"
date: "17/09/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages

```{r load-packages}
library(tidyverse)
```

## Exercise 1: For the alcohol-related hospital admission data per intermediate data zone, plot the data distributions for each health board in 2019.

We can read in the data we saved before, which includes alcohol-related hospital admissions and health boards.

```{r read-admission-healthboard-data}

admission_data <- read_csv(here("INPUTS/WEEK7/alcohol_related_admissions.csv")) 
glimpse(admission_data)

```

## Create histogram of hospital admissions for each health board in 2019

```{r plot-data-distribution}

admission_data %>% 
  filter(year == 2019) %>%
  ggplot(aes(x = alcohol_admissions)) + 
  geom_histogram(bins = 15) +
  facet_wrap(~health_board)

```

## Exercise 2: For which health boards do we have data on alcohol-related hospital admissions for each deprivation quintile? 

First we need to read in the data

```{r read-admission-deprivation-data}
# alcohol-related hospital admissions
data <- read_csv(here("INPUTS/WEEK7/alcohol-admissions_deprivation-data.csv")) # alcohol-related hospital admissions

# intermediate data zone info
intermediate_zone_codes <- read_csv(here("INPUTS/WEEK7/iz2011_codes_and_labels_21042020.csv")) 

```

## Prepare data
We need different information from the geography dataset.

```{r select-columns}
glimpse(data)

data <- data %>%
  select(geography_code,quintile,period,indicator_measure)

glimpse(intermediate_zone_codes)

HB_info <- intermediate_zone_codes %>%
  select(HB,HBName)

```


## Join the datasets
We need to join the datasets so that we get the health board name in the same dataset as the hospital admissions.

Here there are multiple rows that match in the two datasets, so we get duplicate rows in the joined dataset. We remove duplicates using the distinct function.

```{r join-datasets}
admission_data <- left_join(data,HB_info,by = c("geography_code" = "HB")) %>%
  distinct()

glimpse(admission_data)

# Remove "NHS" from the health board name, and "financial year" from the period

admission_data <- admission_data %>%
  mutate(HBName = gsub("NHS ", "", HBName),
         period = gsub(" financial year","", period)) %>%
  rename(health_board = HBName,
         alcohol_admissions = indicator_measure)

glimpse(admission_data)
```

Then, we can find out how many health boards we have data for by using the unique function:

```{r unique-health-boards}

admission_data$health_board %>%
  unique()

```


## Exercise 3: Create boxplots showing alcohol admissions for each deprivation quintile, in a different plot for each health board in the data set. 

Show the data points for each period (using the "jitter" plot).
Remove the "Total" deprivation value first.

```{r boxplots}

admission_data %>%    
  filter(quintile != "Total") %>%
  ggplot(aes(x = quintile,             
             y = alcohol_admissions)) +    
  geom_boxplot() +                        
  geom_jitter(alpha = 0.4) +               
  theme(legend.position = "none") + 
  facet_wrap(~health_board)          

```


