---
title: "Hip Replacements"
author: "DBlana"
date: "17/08/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load packages

```{r load-packages}
library(tidyverse)
```
### Read in the data

```{r read-data}
hip_data <- read_csv(here("INPUTS/WEEK6/Hip Replacement CCG 1819.csv"))
```
### Inspect the data

```{r inspect-data}
glimpse(hip_data)
```

## Exercise 1: Plot 'EQ-5D Index' scores pre and post operation for each gender

### Select gender and quality of life score pre and post operation

```{r select-gender-eq5d}
gender_EQ5D <- hip_data %>%
  select(`Gender`,`Pre-Op Q EQ5D Index`,`Post-Op Q EQ5D Index`) %>%
  rename(EQ5D_Pre = `Pre-Op Q EQ5D Index`,
         EQ5D_Post = `Post-Op Q EQ5D Index`
         )

head(gender_EQ5D)
```

### Identify and remove missing values

```{r missing-values}
gender_EQ5D$Gender %>% unique()
gender_EQ5D$Gender %>% table()
gender_EQ5D %>% summary()

gender_EQ5D_noNA <- gender_EQ5D %>%
  drop_na() %>%
  filter(Gender !='*')

table(gender_EQ5D_noNA$Gender)
summary(gender_EQ5D_noNA)
```
### Replace 1 and 2 with "male" and "female"

```{r recode-gender}
gender_EQ5D_noNA <-gender_EQ5D_noNA %>%
  mutate(Gender = ifelse(Gender == 1, "male", "female"))

```


### Check that data is tidy
The data frame is not tidy, because the column names EQ5D_Pre and EQ5D_Post contain *data*: the time point when EQ5D was measured: pre or post operation.  

```{r tidy-data}
head(gender_EQ5D_noNA)

tidy_gender_EQ5D_noNA <- gender_EQ5D_noNA %>%
  pivot_longer(c(EQ5D_Pre,EQ5D_Post),
               names_to = 'Time',   # the name of the column to create from the data stored in the original column names
               names_prefix = 'EQ5D_',  # remove this text from the start of each variable name
               values_to = 'EQ5D' # the name of the column to create from the data stored in cell values
               )

head(tidy_gender_EQ5D_noNA)
  
```


### Plot quality of life pre and post operation for each gender

```{r plot-eq5d}

# Turn Time into a "factor" so we can order the categories any way we want
# otherwise they are alphabetical and "Post" ends up before "Pre"
tidy_gender_EQ5D_noNA$Time <- factor(tidy_gender_EQ5D_noNA$Time,levels=c('Pre','Post'))

# ggplot creates a blank canvas, to which we add a boxplot with "geom_boxplot"
tidy_gender_EQ5D_noNA %>%
  ggplot() +
  geom_boxplot(aes(x = Time, y = EQ5D, colour = Gender))
```

## Exercise 2: Calculate how many patients in this dataset have been told by a doctor that they have problems caused by a stroke

### Select Stroke column only

```{r select-stroke-data}
stroke_data <- hip_data %>%
  select(Stroke)

head(stroke_data)
```
### Calculate frequencies

```{r stroke-frequencies}
table(stroke_data)
```

## Exercise 3: Create a clean and tidy table with pre and post operation activity levels

### Select activity pre and post operation

```{r select-activity-data}
activity_data <- hip_data %>%
  select(`Pre-Op Q Activity`,`Post-Op Q Activity`) %>%
  rename(Activity_Pre = `Pre-Op Q Activity`,
         Activity_Post = `Post-Op Q Activity`
         )

head(activity_data)
```

### Remove missing values

```{r activity-missing-values}
activity_data %>% table()

activity_data_noNA <- activity_data %>%
  filter(Activity_Pre != 9,
         Activity_Post != 9)

activity_data_noNA %>% table()
```

### Check data is tidy

```{r activity-tidy-data}
head(activity_data_noNA)

tidy_activity_data_noNA <- activity_data_noNA %>%
  pivot_longer(c(Activity_Pre,Activity_Post),
               names_to = 'Time',
               names_prefix = 'Activity_',
               values_to = 'Activity'
               )

head(tidy_activity_data_noNA)
  
```

```
```

