---
title: "Answers to exercises on data manipulation"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load packages
We will be using the readr and dplyr packages from the Tidyverse family of packages. We will also load the "here" package that we will use to read in our data.

```{r}
library(readr)
library(dplyr)
library(here)
```

### FEV dataset

Our data is in file fev.csv, we will save it in an object called fev_data:

```{r}
fev_data <- read_csv(here("INPUTS/WEEK6/fev.csv"))
```

**Problem 1**: Which subjects are male and which are female? (i.e. what does sex == 1 mean?)
From the variables in our dataset, the only one that could be related to sex is height: we might expect that men are generally taller than women.

```{r}
fev_data %>%
  group_by(sex) %>%
  summarise(mean(height), sd(height))
```
This seems to imply that 1 == men. But are men taller than women in a young age? Let's look at the ages included in our dataset:

```{r}
table(fev_data$age)
```
We have people from 3 to 19 years old. Perhaps we should only look at older people:

```{r}
fev_data %>%
  filter(age>15) %>%
  group_by(sex) %>%
  summarise(mean(height), sd(height))
```
Or we could group by sex and age:

```{r}
fev_data %>%
  group_by(age,sex) %>%
  summarise(n(),mean(height), sd(height)) %>%
  arrange(desc(age)) %>%
  print(n=40)
```

It does seem that people with sex == 1 at older ages are taller, so this suggests that 1 == men. 

**Problem 2**: Why do smokers appear to have better lung function (higher forced expiratory volume - FEV)?

Let's check first: is this true?

```{r}
fev_data %>%
  group_by(smoke) %>%
  summarise(mean(fev))
```
It does seem that smokers have higher FEV. But FEV is also dependent on age: younger people have lower FEV. As before, we could group by both smoke and age:  

```{r}
fev_data %>%
  group_by(age,smoke) %>%
  summarise(n = n(), mean = mean(fev), sd = sd(fev)) %>%
  filter(n>5) %>%  # here I am only including rows where we have at least five people - to remove low counts
  arrange(desc(age))
```

Now we can compare smoker and non-smokers at the same age, and we can see the smokers generally have lower FEV, as expected.